Draft Report

### 2020-11-17 DraftReport.Rmd
variable rename update model2, model3

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate') %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         Black...of.Total.Population, White...of.Total.Population,
         Percent.living.under.the.federal.poverty.line..2018., Percent.Unemployed..2018., X65.)
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

@@

 - Children 0-18
 - Adults 19-25
 - Adults 26-34
 - Adults 35-54
 - Adults 55-64
 - 65+
 - white % of Total Population,
 - Black % of Total Population, 
 - Hispanic % of Total Population, 
 - Other % of Total Population

Currently, Case Rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state. Variables such as Population density per square miles, Percent Unemployed (2018)/Percent living under the federal poverty line (2018), White % of of Total Population/Black % of Total Population, Percentage of 65+, and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.

@@

In addition, correlation values of White% and Black %, Percent Unemployed and Percent living under the federal poverty line shows these two groups of variables are correlated respectively.

@@

```{r}
ggplot(data = df,
  mapping = aes(x= case_rate_100k))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "Cases per 100,000", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

@@

```{r}
histogram_of_pdensity <- df %>%
  ggplot(aes(x = population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Distribution of Population Density per Square Mile', 
    x = 'Population density per square mile', y = 'Count')

@@

# Plot the new distribution in a histogram
histogram_of_pdensity <- df %>%
  ggplot(aes(x = l_population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Figure 2. Distribution of Population Density per Square Mile without D.C.', 
    x = 'Log of population density per square mile', y = 'Count')

@@

```{r}
df %>% 
  ggplot(aes(l_population_density, case_rate_100k)) + 
  geom_smooth(se = FALSE) +
  geom_point() +   
  labs(
    title = 'COVID Case Rate due to Population Density', 
    x = 'Log of population density per square mile', 
    y = 'COVID case rate per 100,000'
  )
```

@@

```{r}
black_num <- as.numeric(df$Black...of.Total.Population)
black_num[is.na(black_num)]<- 0
cor(df$White...of.Total.Population, black_num)
```

```{r}
summary(df$White...of.Total.Population)
```

```{r}
ggplot(data = df,
  mapping = aes(x= White...of.Total.Population))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
  labs(title =  "Histogram of White Percentage", x = "White % of Total Population", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 1, 0.1))
```
```{r}
ggplot(data = df,
  mapping = aes(x= (White...of.Total.Population)^2))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
  labs(title =  "Histogram of White Percentage", x = "White % of Total Population^2", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 1, 0.1))
```

#### 2-b) Percent living under the federal poverty line 2018

Correlation of pecent living under poverty line and white percentage of total population shows that they are not significantly related. The distribution is not heavily skewed.

```{r}
summary(df$Percent.living.under.the.federal.poverty.line..2018.)
cor(df$Percent.living.under.the.federal.poverty.line..2018., df$White...of.Total.Population)
```

```{r}
ggplot(data = df,
  mapping = aes(x= Percent.living.under.the.federal.poverty.line..2018.))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
  labs(title =  "Histogram of Percent of poverty 2018", x = "Percent.living.under.the.federal.poverty.line", y = 'Count')+
  scale_x_continuous(breaks=seq(6, 21, 1))
```
#### 2-c) Percent Unemployed 2018
Although, the distribution is of high peak in the middle of the range. Overall, it is not highly skewed or heavily tailed. In addition, the unemployed rate is correlated to the variable to the poverty rate, which is in an agreement with intuitive expectation.

```{r}
summary(df$Percent.Unemployed..2018.)
cor(df$Percent.Unemployed..2018., df$Percent.living.under.the.federal.poverty.line..2018.)
```
```{r}
ggplot(data = df,
  mapping = aes(x= Percent.Unemployed..2018.))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) +
  labs(title =  "Histogram of Unemployed percentage", x = "Percent.Unemployed..2018.", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 10, 1))
```

#### 2-d) Age 65+ Percentage
The 65+ percentage is of a nearly normal distribution.

```{r}
summary(df$X65.)
```
```{r}
ggplot(data = df,
  mapping = aes(x= X65.))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0.10,0.22,0.01)) +
  labs(title =  "Histogram of Percent of 65+", x = "65+ Percentage", y = 'Count')+
  scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
```

@@

```{r}
df %>% 
  ggplot(aes(log10(population_density), case_rate_100k, color = White...of.Total.Population)) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to Population Density', 
    x = 'Log Population Density', 
    y = 'Case.Rate.per.100000',
    color = 'White %'
  )
```
```{r}
df %>% 
  ggplot(aes(x = log10(population_density), y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to Population Density', 
    x = 'log Population Density', 
    y = 'Case.Rate.per.100000',
    color = 'Poverty Percentage'
  )
```

```{r}
df %>% 
  ggplot(aes(x = White...of.Total.Population, y = case_rate_100k, color = Percent.living.under.the.federal.poverty.line..2018.)) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to White %', 
    x = 'White % of Total Population', 
    y = 'Case.Rate.per.100000',
    color = 'Poverty Percentage'
  )
```

```{r}
df %>% 
  ggplot(aes(x = Percent.living.under.the.federal.poverty.line..2018., y = case_rate_100k, color = log10(population_density))) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to Poverty %', 
    x = 'Percent.living.under.the.federal.poverty.line..2018.', 
    y = 'Case.Rate.per.100000',
    color = 'log(Population Density)'
  )
```
```{r}
df %>% 
  ggplot(aes(X65., case_rate_100k, color = log10(population_density))) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to 65+ Percentage', 
    x = 'X65+.', 
    y = 'Case.Rate.per.100000',
    color = 'log(Population density)'
  )
```

```{r}
df %>% 
  ggplot(aes(Percent.Unemployed..2018., case_rate_100k, color = log10(population_density))) + 
  geom_point() +   
  labs(
    title = 'Relation of Case.Rate.per.100000 to Unemployed Percentage', 
    x = 'Percent.Unemployed..2018.', 
    y = 'Case.Rate.per.100000',
    color = 'log(Population density)'
  )
```

@@

#### Model 2-a
$$
\begin{aligned}
 Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.total. population)
\end{aligned}
$$

```{r}
model2a <- lm(case_rate_100k ~ log10(population_density)+ White...of.Total.Population, data = df)
coeftest(model2a, vcov = vcovHC)

@@

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.total. population)^2
\end{aligned}
$$
```{r}
model2a <- lm(case_rate_100k ~ log10(population_density)+ I((White...of.Total.Population)^2), data = df)
coeftest(model2a, vcov = vcovHC)
```

@@

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles) + \beta_3(Percen.living.under.the.federal.poverty.line.2018)
\end{aligned}
$$
```{r}
model2b <- lm(case_rate_100k ~ log10(population_density) + Percent.living.under.the.federal.poverty.line..2018., data = df)
coeftest(model2b, vcov = vcovHC)
```

@@

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0+\beta_1 log(Population.Density.Per.Square.Miles)+\beta_4(Percent.Unemployed..2018.)
\end{aligned}
$$
```{r}
model2c <- lm(case_rate_100k ~ log10(population_density) + Percent.Unemployed..2018., data = df)
coeftest(model2c, vcov = vcovHC)
```

@@

$$
\begin{aligned}
Case.Rate.Per.100000=\beta_0+\beta_1log(Population.Density.Per.Square.Miles)+\beta_4(65+)
\end{aligned}
$$
```{r}
model2d <- lm(case_rate_100k ~log10(population_density)+ X65., data = df)
coeftest(model2d, vcov = vcovHC)
```

@@

```{r}
model3a <- lm(case_rate_100k ~ log10(population_density)
              + Percent.living.under.the.federal.poverty.line..2018.
              + mask_public_bool, data = df)
coeftest(model3a, vcov = vcovHC)
```
#### Model 3-b&c
Considering that there could be multiple policy variables that are relevant to predicting the case number. We have also examined the "Closed.other.non.essential.businesses" variable (renamed as "business_closed") which represents the date when the state issued closure of non-essential businesses. As "business_closed" is a date variable, a transformation was performed to convert the date into a variable "business_close_diff" which represents the days difference between the business close date of the state compared to the average date of all the states.
```{r}

>>>>

```{r variables, warning=FALSE, results='hide'}
df <- read.csv("covid-19.csv", header = TRUE)
df<-df%>%
  rename(case_rate_100k = 'Case.Rate.per.100000',
         death_rate_100k = 'Death.Rate.per.100000',
         population_density = 'Population.density.per.square.miles',
         white_pct = 'White...of.Cases',
         black_pct = 'Black...of.Cases',
         hispanic_pct = 'Hispanic...of.Cases',
         other_pct = 'Other...of.Cases',
         state_emergency = 'State.of.emergency',
         business_closed = 'Closed.other.non.essential.businesses',
         business_reopen = 'Began.to.reopen.businesses.statewide',
         mask_public='Mandate.face.mask.use.by.all.individuals.in.public.spaces',
         mask_legal='No.legal.enforcement.of.face.mask.mandate',
         black_population_pct ="Black...of.Total.Population",
         white_population_pct = "White...of.Total.Population",
         poverty_pct = "Percent.living.under.the.federal.poverty.line..2018.",
         unemployed_pct = "Percent.Unemployed..2018.",
         senior_pct = "X65.") %>%
  select(State, case_rate_100k, death_rate_100k,population_density, white_pct, black_pct, hispanic_pct, 
         other_pct, state_emergency, business_closed, business_reopen, mask_public, mask_legal, 
         black_population_pct, white_population_pct, poverty_pct, unemployed_pct, senior_pct)
# Assign correct data types
cols.num <- c("white_pct","black_pct","hispanic_pct","other_pct")
df[cols.num] <- sapply(df[cols.num],as.numeric)
df$state_emergency=as.Date(df$state_emergency, format = "%m/%d/%Y")
df$business_closed=as.Date(df$business_closed, format = "%m/%d/%Y")
df$business_reopen=as.Date(df$business_reopen, format = "%m/%d/%Y")
head(df)
```

@@

 - Children 0-18 Percentage
 - Adults 19-25 Percentage
 - Adults 26-34 Percentage
 - Adults 35-54 Percentage
 - Adults 55-64 Percentage
 - 65+ Percentage
 - white Percentage of Total Population,
 - Black Percentage of Total Population, 
 - Hispanic Percentage of Total Population, 
 - Other Percentage of Total Population

Currently, Case rate per 100,000 is chosen as a dependent variable to represent the pandemic status in each state. Variables such as population density per square miles, unemployed (2018)/percentage living under the federal poverty line (2018), white/black percentage of of total Population, percentage of seniors(65+), and mask mandate policies are chosen as candidate covariates to explore potential relationships between pandemic status demographic characteristics, economic status, racial composition, and state policy.

@@

In addition, correlation values of white percentage and black percentage, unemployed percentage and percentage living under the federal poverty line shows these two groups of variables are correlated respectively.

@@

```{r}
ggplot(data = df,
  mapping = aes(x= case_rate_100k))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,6000,600)) +
  labs(title =  "Histogram of COVID Case Rate per 100,000", x = "COVID case rate per 100,000", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 6000, 600))
```

@@

```{r}
histogram_of_pdensity <- df %>%
  ggplot(aes(x = population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Distribution of Population Density per Square Miles', 
    x = 'Population density per square mile', y = 'Count')

@@

# Plot the new distribution in a histogram
histogram_of_pdensity <- df %>%
  ggplot(aes(x = l_population_density)) + 
  geom_histogram(fill = 'skyblue', color = 'grey30', bins = 20) + 
  labs(
    title = 'Figure 2. Distribution of Population Density per Square Miles', 
    x = 'Log 10 of population density per square mile', y = 'Count')

@@

```{r}
df %>% 
  ggplot(aes(l_population_density, case_rate_100k)) + 
  geom_smooth(se = FALSE) +
  geom_point() +   
  labs(
    title = 'COVID Case Rate due to Population Density', 
    x = 'Log 10 of population density per square miles', 
    y = 'COVID case rate per 100,000'
  )
```

@@

```{r}
black_num <- as.numeric(df$black_population_pct)
black_num[is.na(black_num)]<- 0
cor(df$white_population_pct, black_num)
```

```{r}
summary(df$white_population_pct)
```

```{r}
ggplot(data = df,
  mapping = aes(x= white_population_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
  labs(title =  "Histogram of White Population Percentage", x = "White percentage of total population", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 1, 0.1))
```
```{r}
ggplot(data = df,
  mapping = aes(x= (white_population_pct)^2))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,1,0.05)) +
  labs(title =  "Histogram of Square of White Percentage", x = "White percentage of total population^2", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 1, 0.1))
```

#### 2-b) Percentage living under the federal poverty line(2018)

Correlation of pecentage living under poverty line and white percentage of total population shows that they are not significantly related. The distribution of the poverty percentage is not heavily skewed.

```{r}
summary(df$poverty_pct)
cor(df$poverty_pct, df$white_population_pct)
```

```{r}
ggplot(data = df,
  mapping = aes(x= poverty_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(6,21,1)) +
  labs(title =  "Histogram of Poverty Percentage", x = "Percentage living under poverty line(2018)", y = 'Count')+
  scale_x_continuous(breaks=seq(6, 21, 1))
```
#### 2-c) Unemployed Percentage (2018)
Although, the distribution is of high peak in the middle of the range. Overall, it is not highly skewed or heavily tailed. In addition, the unemployed rate is correlated to the variable to the poverty rate, which is in an agreement with intuitive expectation.

```{r}
summary(df$unemployed_pct)
cor(df$unemployed_pct, df$poverty_pct)
```
```{r}
ggplot(data = df,
  mapping = aes(x= unemployed_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0,10,1)) +
  labs(title =  "Histogram of Unemployed Percentage", x = "Unemployed percentage (2018)", y = 'Count')+
  scale_x_continuous(breaks=seq(0, 10, 1))
```

#### 2-d) Seniors Percentage(Age 65+)
The seniors percentage (age 65+) is of a nearly normal distribution.

```{r}
summary(df$senior_pct)
```
```{r}
ggplot(data = df,
  mapping = aes(x= senior_pct))+
  geom_histogram(fill = 'skyblue', color = 'grey30', breaks = seq(0.10,0.22,0.01)) +
  labs(title =  "Histogram of Seniors Percentage (65+)", x = "Seniors percentage", y = 'Count')+
  scale_x_continuous(breaks=seq(0.1, 0.25, 0.05))
```

@@

```{r}
df %>% 
  ggplot(aes(white_population_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate Per 100000 to White Population Percentage', 
    x = 'White percentage of total population', 
    y = 'Case rate per 100000',
    color = 'Log 10 of population density'
  )
```
```{r}
df %>% 
  ggplot(aes(x = white_population_pct, y = case_rate_100k, color = poverty_pct)) + 
  geom_point() +
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to White Population Percentage', 
    x = 'White percentage of total population', 
    y = 'Case rate per 100000',
    color = 'Poverty percentage'
  )
```
```{r}
df %>% 
  ggplot(aes(x = poverty_pct , y = case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Poverty Percentage', 
    x = 'Poverty Percentage', 
    y = 'Case rate per 100000',
    color = 'Log 10 of population density'
  )
```

```{r}
df %>% 
  ggplot(aes(unemployed_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() + 
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Unemployed Percentage', 
    x = 'Unemployed percentage(2018)', 
    y = 'Case rate per 100000',
    color = 'log 10 of population density'
  )
```

```{r}
df %>% 
  ggplot(aes(senior_pct, case_rate_100k, color = log10(population_density))) + 
  geom_point() +   
  geom_smooth(se=FALSE)+
  labs(
    title = 'Relation of Case Rate per 100000 to Seniors Percentage(65+)', 
    x = 'Seniors percentage', 
    y = 'Case rate per 100000',
    color = 'log 10 of population density'
  )
```

@@

#### Model 2-a
$$
\begin{aligned}
 Case.Rate.Per.100000 =\beta_0 &+\beta_1 log(Population.Density.Per.Square.Miles)\\&+\beta_2(White.Percentage.of.Total.Population)
\end{aligned}
$$

```{r}
model2a <- lm(case_rate_100k ~ log10(population_density)+ white_population_pct, data = df)
coeftest(model2a, vcov = vcovHC)

@@

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles)\\ &+ \beta_2(White.Percentage.of.Total.Population)^2
\end{aligned}
$$
```{r}
model2a <- lm(case_rate_100k ~ log10(population_density)+ I((white_population_pct)^2), data = df)
coeftest(model2a, vcov = vcovHC)
```

@@

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0 +\beta_1 log(Population.Density.Per.Square.Miles) + \beta_3(Percentage.Under.Poverty.Line.2018)
\end{aligned}
$$
```{r}
model2b <- lm(case_rate_100k ~ log10(population_density) + poverty_pct, data = df)
coeftest(model2b, vcov = vcovHC)
```

@@

$$
\begin{aligned}
 Case.Rate.Per.100000 &=\beta_0+\beta_1 log(Population.Density.Per.Square.Miles)+\beta_4(Percentage.Unemployed.2018.)
\end{aligned}
$$
```{r}
model2c <- lm(case_rate_100k ~ log10(population_density) + unemployed_pct, data = df)
coeftest(model2c, vcov = vcovHC)
```

@@

$$
\begin{aligned}
Case.Rate.Per.100000=\beta_0+\beta_1log(Population.Density.Per.Square.Miles)+\beta_4(Seniors.Percentage)
\end{aligned}
$$
```{r}
model2d <- lm(case_rate_100k ~log10(population_density)+ senior_pct, data = df)
coeftest(model2d, vcov = vcovHC)
```

@@

```{r}
model3a <- lm(case_rate_100k ~ log10(population_density)
              + poverty_pct
              + mask_public_bool, data = df)
coeftest(model3a, vcov = vcovHC)
```
#### Model 3-b&c
Considering that there could be multiple policy variables that are relevant to predicting the case number. We have also examined the "Closed.other.non.essential.businesses" variable (renamed as "business_closed") which represents the date when the state issued closure of non-essential businesses. As "business_closed" is a date variable, a transformation was performed to convert the date into a variable "business_close_diff" which represents the days difference between the business close date of the state compared to the average date of all the states.
```{r}
